### 2020-02-23 manuscript/ERL_global_C_review.Rmd
removed Fig.1, started reworking intro
#19

global C cycle figure files moved to: https://github.com/teixeirak/Global_C_cycle

Forest ecosystems globally influence climate through their critical role in the global carbon (C) cycle (Fig. 1). Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at 59 Gt C yr-1 (Beer et al 2010){**Bagdley et al. is better estimate: >69, also reviews other more recent estimates**}, or 6.3 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017). A small portion of global terrestrial $GPP$ is retained in ecosystems (mainly forests), resulting in a C sink that averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017). The remaining ~98% of global $GPP$ is counterbalanced by ecosystem respiration (Reco) or wildfire. Perturbation to the global $GPP$- Reco balance can substantially influence the growth rate of atmospheric CO~2~; for example, the 2015-2016 El Niño, which brought historically high temperatures and low precipitation to the tropics, released an extra 2.5 ± 0.3 Gt C to the atmosphere, resulting in the largest recorded atmospheric CO~2~ growth rate (Le Quéré et al 2016, Liu et al 2017). In addition, forests contain substantial C stocks that, when disturbed, release significant amounts of CO~2~ to the atmosphere. Although they cover only ~30% of the land surface, forests contain an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008). Globally, gross tropical deforestation averaged 2.8 Gt C yr-1 from 2000-2007, but ~40% of this was offset by forest regrowth, resulting in a net source of ~1.1 Gt C yr-1 from tropical land use change (Pan et al 2011). This, coupled with minimal net deforestation in the extratropics and net uptake by intact forests (Pan et al 2011, Schimel et al 2015), resulted in a total gain in forest C of ~1.2-1.7 Gt C yr-1 (Le Quéré et al 2017, Schimel et al 2015), thereby substantially slowing the rate of increase of atmospheric CO~2~.

![**Figure 1 | The role of forests in the global carbon cycle. Values apply to the period 2000-2018.** Sources are as follows: biomass- (Pan et al 2013, Baccini et al 2012) IPCC 2000; intact forest sink, net forest sink, tropical deforestation and regrowth-(Pan et al 2011); terrestrial sink- (Le Quéré et al 2017); $GPP$- (Beer et al 2010); respiration+fire: calculated here; fossil fuel emissions- (Le Quéré et al 2017); atmospheric CO~2~ (2018 value from https://scripps.ucsd.edu/programs/keelingcurve/). **see issue 19: https://github.com/forc-db/ERL-review/issues/19** ](tables_figures/Global_C_cycle/Global C cycle diagram.png)

Given their vital role in regulating atmospheric CO~2~, the future of Earth’s forests will strongly influence the course of climate change. If forests globally respond to the suite of contemporary and future global change drivers—including elevated CO~2~, climate change, and atmospheric deposition—with increased productivity and net biomass increases, they could continue to act as a significant buffer against anthropogenic emissions. In contrast, if factors such as higher temperatures and drought stress lead to net C losses, the C sink of forests could be lost and even reversed. Global coupled climate models vary substantially in predictions regarding the future of the global forest C sink, with uncertainty in end-of-century terrestrial NEP exceeding current annual anthropogenic CO~2~ emissions (Friedlingstein et al 2006, Cavaleri et al 2015). Anthropogenic land use decisions will also strongly influence the future course of atmospheric CO~2~. In the absence of measures to reduce deforestation, this will continue to be a significant component of anthropogenic CO~2~ emissions. On the other hand, forest conservation, reforestation, and forest management hold strong potential to help stabilize and eventually reduce total anthropogenic CO~2~ emissions (estimated total of 4.4 Gt C yr-1; Griscom et al 2017; *Cook-Patton et al.*; **other refs?**) and thereby contribute towards the goals of the Paris Climate Agreement (UNFCCC 2015, Houghton et al 2015; **REFS**). Indeed, forest-based climate mitigation was a key component of Paris Climate Agreement commitments, totaling approximately one-fourth of nations’ planned emissions reductions (Grassi et al 2017; IPCC1.5). Thus, forest will strongly influence the future course of climate change both through their responses to global change and through forest-related anthropogenic land use decisions.

While forests will play a critical role in the future course of climate change, there remain significant uncertainties as to current and future C cycling the world’s forests. Even as the climate changes and forest age distributions shift, there are important gaps in our fundamental understanding of how C cycling varies with climate and stand age. A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling. Recent studies attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (*i.e.*, dead wood, litter, soil), and source/sink status of forests in some regions. Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012). To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

Here, we synthesize ForC data to provide a big-picture overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad questions:

>>>>

Carbon (C) cycling in forests is central to sustaining life on Earth. Forest ecosystems globally influence climate through their vital role in regulating atmospheric carbon dioxide (CO~2~). Their annual gross CO~2~ sequestration (gross primary productivity, $GPP$) is estimated at >69 Gt C yr^-1^ (Bagdley et al. 2019), or >7 times average annual fossil fuel emissions from 2007-2016 (9.4 ± 0.5 Gt C yr-1; Le Quéré et al 2017) (**update**). While most of this enormous C flux is couterbalanced by CO~2~ releases to the atmosphere through ecosystem respiration ($R_{eco}$) or fire, a small portion was retained in ecosystems over recent decades. The resulting in C sink averaged 3.0 ± 0.8 Gt C yr-1 from 2007-2016, offsetting 32% of anthropogenic fossil fuel emissions (Le Quéré et al 2017) (**update, give range**). Moreover, forests contain substantial C stocks: an estimated 92% of terrestrial biomass (Pan et al 2013) and 45% of terrestrial C (biomass and soils; Bonan 2008). Globally, net deforestation (*i.e.*, gross deforestation - regrowth) has been a source of CO~2~ emissions, estimated at ~1.1 Gt C yr-1 from YEAR-YEAR (Pan et al 2011), resulting in net C sink of ~1.2-1.7 Gt C yr^-1^ across Earth's forests (Le Quéré et al 2017, Schimel et al 2015) (**update, give range**). The future of this sink is dependent both upon forest responses to a broad suite of global change drivers and to future land use decisions, and will strongly influence the course of climate change. Understanding, modeling, and managing forest-atmosphere CO~2~ exchange is thereby central to efforts to mitigate climate change (Friedlingstein et al 2006, Cavaleri et al 2015, Griscom et al 2017; Grassi et al 2017; IPCC1.5).

On an ecosystem level, forest C cycling is central to the flow of energy and materials, and as the basis of food webs...

Despite the centrality of forest C cycling in regulating both atmospheric CO~2~ and ecosystem function and diversity, we lack a comprehensive understanding of how C cycling varies with across biomes and with stand age. While satellite measurements are increasingly useful for estimating forest $GPP$ (Bagdley et al. 2019), aboveground biomass ($B_{ag}$) (REFS), and possibly woody mortality (*i.e.*, $B_{ag}$ losses to mortality $M_{woody}$) (REFS), the remainder of C fluxes and stocks can only be quantified by ground-based measurements. ...

*A relative dearth of data on C cycling in secondary forests, particularly in the tropics (Anderson-Teixeira et al 2016), is problematic in that almost 2/3 of the world’s forests were secondary as of 2010 (FAO 2010), implying an under-filled need to characterize age-related trends in forest C cycling. Recent studies attempting to quantify the role of forests in the global C cycle (*e.g.*, Pan et al 2011) have been hampered by insufficient data on regrowth rates of tropical forests worldwide, C in non-living pools (*i.e.*, dead wood, litter, soil), and source/sink status of forests in some regions. Moreover, as discussed above, model representation of forest responses to global change remain highly uncertain (Friedlingstein et al 2006, Cavaleri et al 2015), with efforts to reduce model uncertainty in part limited by ready availability of appropriate benchmark data (Clark et al 2017, Luo et al 2012).*

To address the need for global-scale analyses of forest C cycling, we have developed an open-access Global Forest Carbon database, ForC (https://github.com/forc-db/ForC; Anderson-Teixeira et al 2016, 2018). ForC contains data on forest ecosystem C stocks and annual fluxes (>50 variables) and associated data required for interpretation (*e.g.*, stand history, measurement methods) amalgamated from numerous previous data compilations and directly from original publications. ForC currently contains **#** records from **#** plots and **#** distinct geographic areas representing all forested biogeographic and climate zones.

Here, we synthesize ForC data to provide a macroscopic overview of stand-level carbon cycling of the world’s major forest biomes and how it varies with stand age. We address two broad questions:
